---
title: "IS607 Project1"
author: "Sharon Morris"
date: "September 18, 2016"
output: html_document
---

**Introduction**
In this project, you’re given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:

Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents

For the first player, the information would be:

Gary Hua, ON, 6.0, 1794, 1605

1605 was calculated by using the pre-tournament opponents’ ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and dividing by the total number of games played.

The chess rating system (invented by a Minnesota statistician named Arpad Elo) has been used in many other contexts, including assessing relative strength of employment candidates by human resource departments.

**Import text file into R**

```{r}
tourinfo <- read.table("tournamentinfo.txt", sep = "\n", header=TRUE)

```

**Examin Data**

```{r}
# Top of data
head(tourinfo, 10)

# End of data
tail(tourinfo, 10)

# Structure of the data
str(tourinfo)

# Dimensions
dim(tourinfo)

# Missing values
is.na(tourinfo)

# Data class
class(tourinfo)

```

**Extract player name**

```{r}
library(stringr)
player_name <- (str_trim(unlist(str_extract_all(unlist(tourinfo), 
                "([[:alpha:] ]-?){13,50}"))))[1:64]

head(player_name, 10)
```

**Extrat state**

```{r}
state <- str_trim(unlist(str_extract_all(unlist(tourinfo), "MI | ON | OH ")))

head(state, 10)
```

**Extract pre-rating**

```{r}

# Extact players pre-rating
# check class
class(1794)

pre_ratng <- str_replace_all(str_trim(unlist(str_extract_all(unlist(tourinfo),
                                "R: [[:digit:] ]*"))), "R: ", "")
head(pre_ratng, 10)
```

**Extract round**

```{r}
r <- str_sub(unlist(tourinfo), start = 48, end = 89) [c(seq(1, length(unlist(tourinfo)),
                        by = 3))]
head(r, 10)
```

**Convert data frame to a numeric matrix**
Data are numeric so no need to coerce 

```{r}
# Extract the characters W B D L H
rating <- str_extract_all(r, "( |\\d){4}")
           
head(rating, 10)
```

# Coerce character variables to numeric inorder to convert to numeric matrix

```{r}
rating <- as.numeric(unlist(rating))
```

```{r}
# Convert data frame to numberic matrix
ratng_convert <- matrix (rating, nrow = 1, ncol = 64)
head(ratng_convert, 10)
```

```{r}
# Calculate average
average <- colMeans((ratng_convert))
```

**Print data**

```{r}
data.frame(player_name, state, pre_ratng, average)

```


**Output file**

```{r}
touraverage <- data.frame(player_name, state, pre_ratng, average)
write.csv(touraverage, file="/users/sharonmorris/IS607/tournamentinfodone.txt")
```

