---
title: "Tidying and Transforming Data"
author: "Sharon Morris"
date: "September 25, 2016"
output: pdf_document
---

(1) Create a CSV file (or optionally, a MySQL database!) that includes all of the information above. You’re encouraged to use a “wide” structure similar to how the information appears above, so that you can practice tidying and transformations as described below.

(2)Read the information from your CSV file into R,and use tidyr and dplyr as needed to tidy and transform your data.

(3) Perform analysis to compare the arrival delays for the two airlines.

Import csv file

```{r}
week6 <- read.csv('week5.csv',header=TRUE,stringsAsFactors = FALSE)
week6
```

Remove all rows and missing data

```{r}
library(tidyr)

na <- na.omit(week6)
head(na)
```

Add airline name to rows

Without the added row names the spread will not run because values in rows without names
are treated as duplicates

```{r}

na[2,1] <-  na[1,1] 
na[4,1] <-  na[3,1]

head(na)

```


Insert column names

```{r}

names(na)[1:2] <- c('Airline', 'Status')
na
```

Transform data -- make city into 1 column

```{r}

na <- gather(na, "City", Count,3:7)
na
```

Remove . from state names

```{r}
na$City<-gsub("\\."," ",na$City)

na
```


Separate values in Status column

```{r}
na<-spread(na, Status, Count)

head(na)
```

Rename column on time 

```{r}
colnames(na)[4] <- "Ontime"

head(na)
```

Analysis of airline arrival and departures

```{r}
library(dplyr)
Compare <- na %>% group_by(City, Airline) %>% summarise(ComOntime=sum(Ontime), 
                                        ComDelay=sum(delayed), 
                                   Compariosn=ComOntime/(ComOntime+ComDelay)*100)
Compare
```

